On the problem of repeated supervised learning

### 2023-03-21 paper/main.tex
adding theory and fixing experiment

\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}         % Can be removed after putting your text content
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\newtheorem{theorem}{Theorem}

@@

@@

        In this paper we consider linear model, that is solved as Ridge with mean squared error loss function. In the formulation of our experiment, we consider $\mathbf{R}$ \eqref{R} as space of density functions of random vectors $(\mathbf{x^i}, y_i)$, where $\mathbf{x^i} \in \textbf{X}$ and $y_i \in \textbf{y}$ -- the features and target variable of the $i$-th object. The operator $\text{D}$ transforms $\mathbf{R}$ at each step $t$, and the distribution of features does not change, because at each step we take new $\mathbf{x^i}_t$ from the original set of features $\textbf{X}$, so only the distribution of the target variable changes. 
        If in some step the model starts to give good predictions on the training and test samples, then, in the probabilistic formulation, this means that the density function of the distribution of the object-sign vectors becomes similar to the delta function, as the components of the random vector $(\mathbf{x^i}, y_i)$ become linearly dependent.

>>>>

\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}         % Can be removed after putting your text content
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsthm}
\usepackage{amssymb}
\newtheorem{theorem}{Theorem}

@@

    In this section we will provide several theorems on operator $\text{D}$ to answer questions 1-3.

    \textbf{Notation:} In this paper we will use the common notations: the $L_1(\mathbb{R}^n)$-norm of function $f$:

    \begin{equation*}
        \|f\|_1 := \int\limits_{\mathbb{R}^n} |f(x)| dx \quad \text{ and } \quad L_1(\mathbb{R}^n) := \left\{f : \mathbb{R}^n \to \mathbb{R} ~|~ \|f\|_1 < + \infty\right\}
    \end{equation*}

    The $L_{\infty}(\mathbb{R}^n)$-norm of function $f$:

    \begin{equation*}
        \|f\|_{\infty} := \underset{x \in \mathbb{R}^n}{\text{esssup}}\{f(x)\} := \inf\{C \geq 0 ~|~ |f(x)| \leq C \text{ for almost every } x \in \mathbb{R}^n\} \quad \text{ and } \quad L_{\infty}(\mathbb{R}^n) := \left\{f : \mathbb{R}^n \to \mathbb{R} ~|~ \|f\|_{\infty} < + \infty\right\}
    \end{equation*}

    \begin{theorem}[Fact] \label{Th1}
        If $f: \mathbb{R}^n \to \mathbb{R}$ such that $f(x) \geq 0$ for almost every $x \in \mathbb{R}^n$ and $\|f\|_1 = \int\limits_{\mathbb{R}^n} f(x) dx = 1$, then there exists a random vector $\mathbf{\xi}$, for which $f$ will be a density distribution function.
    \end{theorem}

    Exactly on the basis of Theorem \ref{Th1} we define $\mathbf{R}$ \eqref{R} in this way.

    \begin{theorem}[Assumptions for $\text{D}: \mathbf{R} \to \mathbf{R}$] \label{Th2}
        If $\|\text{D}\|_1 = 1, \forall f \in \mathbf{R} \hookrightarrow \text{D}(f)(x) \geq 0$ for almost every $x \in \mathbb{R}^n$, and exists $\text{D}^{-1}$ such that $\|\text{D}^{-1}\|_1 \leq 1$, then $\text{D} : \mathbf{R} \to \mathbf{R}$.
    \end{theorem}
    \begin{proof}
        To begin with, let us note that if $\text{D}: \mathbf{R} \to \mathbf{R}$, then $\|\text{D}\|_1 = 1$, because by definition of operator norm:

        \begin{equation*}
            \|\text{D}\|_1 \overset{def}{=} \underset{\|f\|_1 = 1}{\sup}\left\{\|\text{D}(f)\|_1\right\}
        \end{equation*}

        And if $f$ such that $\|f\|_1 = 1$ then $|f| \in \mathbf{R}$ and, because $\text{D}: \mathbf{R} \to \mathbf{R}$, $\|\text{D}(|f|)\|_1 = 1$. But $\|\text{D}\|_1 = 1$ is only a necessary but not a sufficient condition.

        If $\|\text{D}\|_1 = 1$, then $\forall f \in \mathbf{R} \hookrightarrow \text{D}(f) \leq 1$. If $\exists f_0 \in \mathbf{R}$ such that $\|\text{D}(f_0)\|_1 < 1$, then we get a contradiction because

        \begin{equation*}
            \|D^{-1}\|_1 \overset{def}{=} \underset{\|f\|_1 \neq 0}{\sup}\left\{\dfrac{\|\text{D}^{-1}(f)\|_1}{\|f\|_1}\right\} 
            \geq \left[f_1 = \text{D}(f_0)\right] \geq
            \dfrac{\|\text{D}^{-1}(f_1)\|_1}{\|f_1\|_1} = 
            \dfrac{\|\text{D}^{-1}(\text{D}(f_0))\|_1}{\|\text{D}(f_0)\|_1} = 
            \dfrac{\|f_0\|_1}{\|\text{D}(f_0)\|_1} = \dfrac{1}{\|\text{D}(f_0)\|_1} > 1
        \end{equation*}

        But we assume that $\|\text{D}^{-1}\|_1 \leq 1$. So $\forall f \in \mathbf{R} \hookrightarrow \|\text{D}(f)\|_1 = 1$.

        But according to Theorem \ref{Th1} to $\text{D} : \mathbf{R} \to \mathbf{R}$ we also need second assumption: $\forall f \in \mathbf{R} \hookrightarrow \text{D}(f)(x) \geq 0$ for almost every $x \in \mathbb{R}^n$.

    \end{proof}

    In our experiment it often difficult to calculate $\text{D}^{-1}$ and especially it's norm, so we make a different assumptions. We assume that our data sample describes the distribution function sufficiently well, so approximating the density function by our sample closely enough approaches the true density function of the data distribution. And if we consider $\text{D}$ as algorithm for training a model, forming predictions on a test sample and mixing predictions into a training sample, then $\text{D} : \mathbf{R} \to \mathbf{R}$ by constructing our density functions of data distributions.

    % Theorem 3 is wrong 
    \if 0
    \begin{theorem}[Inequality on $\|\text{D}\|_{\infty}$] \label{Th3}
        If $\text{D} : \mathbf{R} \cap L_{\infty}(\mathbb{R}^n) \to \mathbf{R} \cap L_{\infty}(\mathbb{R}^n)$ and , then $\|\text{D}\|_{\infty} \geq 1$
    \end{theorem}

    \begin{proof}
        Let us first write out two Herder's inequality: $\forall f \in \mathbf{R} \cap L_{\infty}(\mathbb{R}^n)$ it is fulfilled that

        \begin{equation*}
            \|f\|_1 \cdot \|\text{D}(f)\|_{\infty} \geq \|f \cdot \text{D}(f)\|_1
        \end{equation*}
        \begin{equation*}
            \|f\|_{\infty} \cdot \|\text{D}(f)\|_1 \geq \|f \cdot \text{D}(f)\|_1
        \end{equation*}

        Note that $\|f\|_1 = \|\text{D}(f)\|_1 = 1$ and $f, \text{D}(f) \in \mathbf{R} \cap L_{\infty}(\mathbb{R}^n)$ because $\text{D} : \mathbf{R} \cap L_{\infty}(\mathbb{R}^n) \to \mathbf{R} \cap L_{\infty}(\mathbb{R}^n)$. Let's add these two inequalities together and get

        \begin{equation*}
            \|\text{D}(f)\|_{\infty} + \|f\|_{\infty} \geq 2 \|f \cdot \text{D}(f)\|_1
        \end{equation*}

        Using common inequality on operators norm $\|\text{D}(f)\| \leq \|\text{D}\| \cdot \|f\| ~\forall f \in \mathbf{R}$ we get

        \begin{equation} \label{tmp1}
            (1 + \|\text{D}\|_{\infty}) \|f\|_{\infty} \geq 2 \|f \cdot \text{D}(f)\|_1
        \end{equation}

        Inequality \eqref{tmp1} is fulfilled for any $f \in \mathbf{R} \cap L_{\infty}(\mathbb{R}^n)$. Let's consider for any $\delta > 0$

        \begin{equation*}
            f_{\delta}(x) := V_n(\delta) \cdot \textbf{1}_{B(1/\delta)}(x),
        \end{equation*}

        where $B(1/\delta) := \mathbb{B}^n_{\|\cdot\|_2}(1/\delta) = \{x \in \mathbb{R}^n ~|~ \|x\|_2 \leq 1/\delta$\} -- the euclidean ball with radius $1/\delta$, $\textbf{1}_{A}(x)$ -- characteristic function of a set $A$ and $V_n(\delta) = \frac{\pi^{n/2}}{\Gamma(n/2+1)} \cdot \delta^n$ -- the volume of $\mathbb{B}^n_{\|\cdot\|_2}(\delta)$. In the formula for the volume of an $n$-dimensional Euclidean ball we are interested only in the fact that $V_n(\delta) = \kappa(n) \cdot \delta^n$, so we can calculate

        \begin{equation*}
            \|f_{\delta}\|_1 = \int\limits_{\mathbb{R}^n} |f_{\delta}(x)|dx = V_n(\delta) \cdot \int\limits_{B(1/\delta)}dx = V_n(\delta) \cdot V_n(1/\delta) = 1
            \quad \text{ and } \quad \|f_{\delta}\|_{\infty} = V_n(\delta)
        \end{equation*}

        So, $\forall \delta > 0 \hookrightarrow f_{\delta} \in \mathbf{R} \cap L_{\infty}(\mathbb{R}^n)$ and for it the inequality \eqref{tmp1} is fulfilled:

        \begin{equation*}
            (1 + \|\text{D}\|_{\infty}) \cdot \|f_{\delta}\|_{\infty} \geq 2 \|f_{\delta} \cdot \text{D}(f_{\delta})\|_1
        \end{equation*}

        Assume a notation $g_{\delta} := \text{D}(f_{\delta}) \in \mathbf{R} \cap L_{\infty}(\mathbb{R}^n)$ we obtain

        \begin{equation*}
            (1 + \|\text{D}\|_{\infty}) \cdot V_n(\delta) \geq 2 V_n(\delta) \cdot \|g_{\delta} \cdot \textbf{1}_{B(1/\delta)}\|_1 
        \end{equation*}

        So,

        \begin{equation*}
            1 + \|\text{D}\|_{\infty} \geq 2 \cdot \|g_{\delta} \cdot \textbf{1}_{B(1/\delta)}\|_1
        \end{equation*}

        $g_{\delta} \in \mathbf{R}$, so $\forall \varepsilon > 0 ~~ \exists \delta > 0 : \|g_{\delta} \cdot \textbf{1}_{B(1/\delta)}\|_1 \geq 1 - \varepsilon$, so we obtain

        \begin{equation*}
            \forall \varepsilon > 0 \hookrightarrow 1 + \|\text{D}\|_{\infty} \geq 2 - 2 \varepsilon
        \end{equation*}

        Finally we get $\|\text{D}\|_{\infty} \geq 1$.        
    \end{proof}

    Assumption that $\text{D} : \mathbf{R} \cap L_{\infty}(\mathbb{R}^n) \to \mathbf{R} \cap L_{\infty}(\mathbb{R}^n)$ means that operator $\text{D}$ does not translate <<good>> distribution density functions, for example, into a delta function or some other function that does not lie in $L_{\infty}(\mathbb{R}^n)$.

    Fact that $\|\text{D}\|_{\infty} \geq 1$ means that there always will be function $f \in \mathbf{R}$ such that $\|\text{D}(f)\|_{\infty} \geq \|f\|_{\infty}$, so $\text{D}$ could never be contraction mapping on $\mathbf{R}$ in $\|\cdot\|_{\infty}$, but sequence $\{\text{D}^t(f_0)\}_{t \in \mathbb{N}_0}$ could still be decreasing by $L_{\infty}$-norm, i.e.

    \begin{equation} \label{derc_inf_norm}
        \exists \lambda \in (0; 1) : \forall t \in \mathbb{N} \hookrightarrow 
        \|\text{D}^t(f_0)\|_{\infty} \leq \lambda \|\text{D}^{t-1}(f_0)\|_{\infty}
    \end{equation}

    \fi

    \begin{theorem}[limit in a weak sense to $\delta$ function] \label{delta}
        If $f_t : \mathbb{R} \to \mathbb{R}$ such that $\forall t \in \mathbb{N} \hookrightarrow  \|f_t\|_1 = 1$, $f_t(x) \geq 0$ in almost every point $x \in \mathbb{R}$ and
        \begin{equation} \label{psi}
            \exists \psi : \mathbb{N} \to \mathbb{R} ~\text{ such that }~ \psi \uparrow +\infty ~\text{ i.e. }~ \psi(t+1) \geq \psi(t) ~\forall t \in \mathbb{N} ~\text{ and }~
            \psi(t) \underset{t \to +\infty}{\longrightarrow} +\infty 
        \end{equation}

        And 
        \begin{equation} \label{g}
            \exists g \in L_1(\mathbb{R}) ~\text{ such that }~ \forall t \in \mathbb{N} ~\forall y \in \mathbb{R} \hookrightarrow f_t\left(\dfrac{y}{\psi(t)}\right) \leq \psi(t) \cdot |g(y)|
        \end{equation}

        Then $f_t(x) \underset{t \to \infty}{\longrightarrow} \delta(x)$ in a weak sense, i.e.

        \begin{equation}
            \underset{t \to +\infty}{\lim}\left(\int\limits_{-\infty}^{+\infty} f_t(x) \phi(x) dx\right) = \phi(0),
        \end{equation}

        where $\phi$ is continuous function with compact support
    \end{theorem}

    \begin{proof}
        Assume a notation $I_t = \int\limits_{-\infty}^{+\infty} f_t(x) \phi(x) dx$. Then it's fulfilled that

        \begin{equation*}
            I_t - \phi(0) = \int\limits_{-\infty}^{+\infty} f_t(x) \phi(x) dx - \phi(0) \cdot \int\limits_{-\infty}^{+\infty} f_t(x) dx = \int\limits_{-\infty}^{+\infty} f_t(x) \cdot [\phi(x) - \phi(0)] dx
        \end{equation*}

        The first equation is fulfilled because $\|f_t\| = 1$.

        \newpage

        Replacing the variable $y = \psi(t) \cdot x, dy = \psi(t) \cdot dx$ we get

        \begin{equation} \label{tmp2}
            I_t - \phi(0) = \dfrac{1}{\psi(t)} \cdot \int\limits_{-\infty}^{+\infty} f_t\left(\frac{y}{\psi(t)}\right) \cdot \left[\phi\left(\frac{y}{\psi(t)}\right) - \phi(0)\right] dy
        \end{equation}

        Split the integral \eqref{tmp2} into 3 parts:

        \begin{equation*}
            I_1 := \dfrac{1}{\psi(t)} \cdot \int\limits_{-\infty}^{-A} f_t\left(\frac{y}{\psi(t)}\right) \cdot \left[\phi\left(\frac{y}{\psi(t)}\right) - \phi(0)\right] dy
        \end{equation*}
        \begin{equation*}
            I_2 := \dfrac{1}{\psi(t)} \cdot \int\limits_{-A}^{A} f_t\left(\frac{y}{\psi(t)}\right) \cdot \left[\phi\left(\frac{y}{\psi(t)}\right) - \phi(0)\right] dy
        \end{equation*}
        \begin{equation*}
            I_3 := \dfrac{1}{\psi(t)} \cdot \int\limits_{A}^{+\infty} f_t\left(\frac{y}{\psi(t)}\right) \cdot \left[\phi\left(\frac{y}{\psi(t)}\right) - \phi(0)\right] dy
        \end{equation*}

        Consider the integrals $I_1$. $\phi$ is continuous function with compact support, so $\phi$ is bounded by some constant $M$, i.e. $\forall x \in \mathbb{R} \hookrightarrow |\phi(x)| \leq M$, so

        \begin{equation*}
            \left|I_1\right| \leq \int\limits_{-\infty}^{-A} 2 M \cdot \dfrac{1}{\psi(t)} f_t\left(\frac{y}{\psi(t)}\right) dy \leq \left[\eqref{g}\right]
            \leq \int\limits_{-\infty}^{-A} 2 M \cdot |g(y)| dy 
        \end{equation*}

        Since $g \in L_1(\mathbb{R})$, there exists some constant $A > 0$ such that $\left|I_1\right| \leq \varepsilon$. Similarly, it can be shown that $\left|I_1\right| \leq \varepsilon$.

        Consider the integral $I_2$. $\phi$ is continuous and $\psi \uparrow +\infty$ \eqref{psi}, so $\exists~ T \in \mathbb{N}$ such that $\forall y \in [-A; A]$ $\forall t \geq T$ it is fulfilled that

        \begin{equation*}
            \left|\frac{y}{\psi(t)} - 0\right| \leq \delta ~\text{ and so }~ \left|\phi\left(\frac{y}{\psi(t)}\right) - \phi(0)\right| \leq \varepsilon
        \end{equation*}

        So

        \begin{equation*}
            \left|I_2\right| \leq \varepsilon \cdot \int\limits_{-A}^{A} \dfrac{1}{\psi(t)} f_t\left(\frac{y}{\psi(t)}\right) dy = 
            \varepsilon \cdot \int\limits_{-A}^{A} f_t\left(x\right) dx \leq \varepsilon \cdot \int\limits_{-\infty}^{+\infty} f_t\left(x\right) dx = \varepsilon
        \end{equation*}

        Finally we get

        \begin{equation*}
            \forall \varepsilon > 0 ~~ \exists~ T \in \mathbb{N} : \forall t \geq T \hookrightarrow \left|I_t - \phi(0)\right| \leq 
            \left|I_1\right| + \left|I_2\right| + \left|I_3\right| \leq 3\varepsilon
        \end{equation*}

        So $f_t(x) \underset{t \to \infty}{\longrightarrow} \delta(x)$ in a weak sense.

    \end{proof}

    Let's analyze results of this theorem. Assume that distribution of our data does not change with $t$, i.e. operator $\text{D}$ changes only the distribution of the target variable $y$. If in some step $t$ the model starts to give good predictions on the training and test samples, then, in the probabilistic formulation, this means that the density function of the distribution of the object-sign vectors becomes similar to the delta function, as the components of the random vector $(\mathbf{x^i}, y_i)$ become linearly dependent. So, if density distribution function of $\mathbf{y} - \mathbf{y_{\text{pred}}}$ becomes $\delta$ function.

    \newpage

    Let's analyze formula \eqref{g}. If we take $x = \phi(x)$ then \eqref{g} takes form

    \begin{equation} \label{g_1}
            \exists g \in L_1(\mathbb{R}) ~\text{ such that }~ \forall t \in \mathbb{N} ~\forall x \in \mathbb{R} \hookrightarrow f_t\left(x\right) \leq \psi(t) \cdot |g(x \cdot \psi(t))|
    \end{equation}

    If $x \neq 0$ then $f_t(x) \underset{t \to \infty}{\longrightarrow} 0$, because $g_1(x) := \psi(t) \cdot |g(x \cdot \psi(t))| \in L_1(\mathbb{R})$ since

    \begin{equation*}
        \int\limits_{-\infty}^{+\infty} g_1(x) dx = \int\limits_{-\infty}^{+\infty} \psi(t) \cdot |g(x \cdot \psi(t))| dx = \int\limits_{-\infty}^{+\infty} g_1(z) dz < +\infty
    \end{equation*}

    And so, if $t \to \infty$, then $z := x \cdot \psi(t) \to +\infty$ and $g_1(z) \to 0$, because $g_1 \in L_1(\mathbb{R})$. So, if $x \neq 0$ then $f_t(x) \underset{t \to \infty}{\longrightarrow} 0$.

    Since $\forall t \in \mathbb{R} \hookrightarrow \|f_t\|_1 = 1$, then $f_t(0) \uparrow +\infty$.

    If we substitute $x = 0$ in the \eqref{g_1} then we get $f_t(0) \leq \psi(t) \cdot |g(0)|$, so we can take

    \begin{equation}
        \psi(t) = \dfrac{f_t(0)}{|g(0)|}
    \end{equation}

    Important example of operator $\text{D}$ that translates any function from $\mathbb{R}$ into a $\delta$ function is as follows

    \begin{equation} \label{example1}
        \text{D}^t(f_0)(x) = t \cdot f_0(t \cdot x)
    \end{equation}

    Here we take $g(x) = f_0(x)$ and $\psi(t) = t$. Let's look how this operator translates density functions of normal distribution $\mathcal{N}(0, 5)$ and continuous uniform distribution $\mathcal{U}[-2.5, 2.5]$:

    \begin{figure}[h!]
        \centering
        \includegraphics[width = 0.99\linewidth]{pictures/example1.png}
        \caption{Illustration of weak limit to $\delta$ function. $\mathcal{N}(0, 5)$ left, $\mathcal{U}[-2.5, 2.5]$ right.}
        \label{example1}
    \end{figure}

@@

        In this paper we consider linear model, that is solved as Ridge with mean squared error loss function. In the formulation of our experiment, we consider $\mathbf{R}$ \eqref{R} as space of density functions of random vectors $(\mathbf{x^i}, y_i)$, where $\mathbf{x^i} \in \textbf{X}$ and $y_i \in \textbf{y}$ -- the features and target variable of the $i$-th object. The operator $\text{D}$ transforms $\mathbf{R}$ at each step $t$, and the distribution of features does not change, because at each step we take new $\mathbf{x^i}_t$ from the original set of features $\textbf{X}$, so only the distribution of the target variable changes. So we can use results from Theorem \ref{delta}.
